Image Batching: Batch Ordering Advanced Imaging Tests in the Emergency Department

### 2024-09-09 MgmtSci/MgmtSci.tex
getting going

---
journal: mnsc
title: "Image Batching: Batch Ordering Advanced Imaging Tests in the Emergency Department"
shorttitle: "Image Batching"
authors:
- familyname: Jameson
  othernames: Jacob
  affiliation: Harvard Kennedy School, Harvard University
  email: jacobjameson@g.harvard.edu
- familyname: Soroush
  othernames: Saghafian
  affiliation: Harvard Kennedy School, Harvard University
- familyname: Huckman
  othernames:  Robert
  affiliation: Harvard Business School, Harvard University
- familyname: Hodgson
  othernames:  Nicole
  affiliation: Mayo Clinic, Emergency Department
keywords:
  - 3 to 6 keywords
  - that do not appear in the title
abstract: |
  Abstract goes here
acknowledgment: |
  The authors would like to thank ...
bibliography: references.bib
blind: false
output: rticles::informs_article
header-includes:
  - \usepackage{booktabs}
  - \usepackage{tabularx}
  - \usepackage{graphicx}
  - \usepackage{tikz}
  - \usepackage{siunitx}
  - \usepackage{tablefootnote}
  - \usepackage{longtable}
  - \usepackage{threeparttable}
  - \usepackage{natbib}
  - \usepackage{caption}
  - \usepackage{adjustbox}
  - \usepackage{multirow}
---

# Introduction

Emergency departments (EDs) are high-pressure environments which globally face escalating challenges such as overcrowding, resource limitations, and increasing patient volumes [@mostafa2024strategies, @Sorup2013evaluation]. These challenges can often be exacerbated by the need to provide timely and accurate diagnoses to patients, which often requires the use of advanced imaging tests such as computed tomography (CT) scans, ultrasound, X-ray, and magnetic resonance imaging (MRI) [@waheed2022impact]. The management of diagnostic testing is a critical component of operational performance in EDs, where timely and accurate diagnoses can significantly impact patient outcomes [@committee2015improving]. In many EDs, however, physicians have considerable discretion over how they order these tests, particularly with regard to advanced imaging [@valtchinov2019use]. Yet little is known about the factors driving physicians' decisions to exercise such discretion and how test ordering should be managed when this discretion exists. In this paper, we explore the causal effects of batch ordering advanced imaging tests, a practice in which physicians order multiple tests simultaneously, on operational performance and patient outcomes in the ED.

Worker discretion can improve system performance, but can sometimes enable workers to "choose the 'wrong' task (operationally)" [@vanDonselaar2010ordering]. We consider the physician's decision to order imaging tests for their patient as an optimization problem, where the physician must balance the trade-offs between the benefits of ordering multiple tests simultaneously at the start of the patient encounter, which may expedite the diagnostic process if the tests are necessary for diagnosis and disposition, and the costs of ordering multiple tests simultaneously, which may increase the total time spent in the ED and the resource utilization of the department. Batching stands in contrast to the more common practice of ordering tests sequentially, where physicians order one test at a time, review the results, and then decide whether to order additional tests. The decision to batch order imaging tests is a form of discretion that physicians can exercise in the ED.

@@

We investigate these questions using operational data from two large EDs in the United States. We focus on imaging tests, which are commonly ordered in the ED and are associated with significant resource utilization. We use a unique dataset that contains detailed information on the timing of test orders, the timing of test results, and the timing of patient disposition. We exploit the random assignment of patients to physicians to identify the causal effects of batch ordering imaging tests on operational performance and patient outcomes. We find that physicians deviate from the standard test ordering strategy 42% of the time. We show that physicians are more likely to batch order imaging tests when they are more experienced, when there is an opportunity to follow a superior test ordering strategy by deviating, and when there is an opportunity to batch by deviating. When physicians deviate from the standard test ordering strategy, their average time to disposition increases by 13%. Other performance dimensions, including resource utilization and patient outcomes, are mostly unaffected. Our calculations suggest that forgoing deviations would have led to faster time to disposition, which could have saved 2,494 hours per year, increasing annual profits by 3%.

@@

# Related Literature

# Batch Ordering Advanced Imaging

## Drivers of Image Batching

## Implications for Emergency Department Operations and Patient Outcomes

## Hetereogeneity in Orderging Strategy

# Setting, Data, and Models

## Empirical Setting

@@

## Data

Our primary dataset comes from the ED of the Mayo Clinic of Arizona, a tertiary care hospital without obstetrical services, an inpatient pediatrics unit, or a trauma designation. During the study period from October 6, 2018, through December 31, 2019, the ED recorded 48,854 visits per year, managed across 26 treatment rooms and up to 9 hallway spaces. The department is exclusively staffed by board-eligible or board-certified emergency physicians (EPs), with rotating residents overseeing about 10% of patient volume. The data is summarized in Table \ref{tab:summary_stats}.

@@

A key feature of the Mayo Clinic ED is its rotational patient assignment system, in which patients arriving at the Mayo Clinic ED are randomly assigned to physicians via a rotational patient assignment algorithm @traub2016emergency, which removes potential selection bias concerns for our analyses. In essence, barring arrival time and shift-level variation, the physician-to-patient matching can be deemed random. Table \ref{tab:wald_test} displays that patient encounters (regarding chief complaints and emergency severity) are equitably distributed across physicians within our study's cohort.

@@

Our research design focuses on adults who visit the Mayo Clinic of Arizona ED. We observe $48,854$ such visits during the study period. To improve power, we drop encounters with rare "reasons for visit" (RVF) ($<1000$ total encounters of this kind) and complaints where a batch order occurs less than $5\%$ of the time. Since batch orders are rare for these cases, our physician batch tendency instrument could suffer from a weak instrument problem if we included them. Examples of complaints dropped include Skin Complaints and Urinary Complaints, as well as other complaints where multiple modalities of imaging is unlikely to occur. Excluding these conditions does not introduce selection bias unless physician test batching tendency is orthogonal to physician diagnosing behavior. While this assumption may be violated if we were to use a very detailed level of chief complaint information upon which to base our exclusion criterion, it is plausibly satisfied when using broad complaint categories as we do. In order to estimate a precise measure of physician-level batch tendency, we further restrict our sample to the $15,821$ encounters involving full-time physicians who treat over 500 ED cases per year.

### Instrumental Variable

\
Our explanatory variable in the IV analysis, $Batched_i$, is an indicator for whether patient $i$ has their tests batch ordered during their ED encounter. We define "batching" in line with standard emergency medicine practices. Batching occurs when a physician simultaneously orders a comprehensive set of diagnostic tests, typically covering a broad range of potential diagnoses. This contrasts with sequential ordering, where tests are ordered sequentially based on the information obtained from each test as needed.

@@

### Dependent Variables

\
\textit{Logarithm of Emergency Department Length of Stay (LOS)}: ED LOS is a key performance metric for ED operations. It is defined as the time from patient arrival to the ED to the time of discharge from the ED. We use this metric to evaluate the impact of batch ordering on ED efficiency. This variable is measured in minutes and in our sample is right-skewed. We log-transform this variable to normalize its distribution to meet the assumptions of linear regression.

@@

\textit{Total Imaging Tests Ordered}: Total imaging tests ordered (NUM_IMAGES) is a count of the number of distinct imaging tests ordered for a patient during the ED encounter. We use this variable to evaluate the impact of batch ordering on the number of imaging tests ordered.

@@

## Identification Strategy

Our empirical strategy closely follows the literature that relies on quasi-random assignment of agents to cases, often referred to as the “judges design.” Papers in this literature typically exploit variation in the sentencing leniency of judges who work in the same court. Similarly, we explore batching variation across physicians who work in the same emergency department. In its reduced form, under the assumption of quasi-random assignment, this approach allows researchers to identify the causal effect of being assigned to different types of physicians (i.e. physiancs with a higher or lower tendency to batch order imaging). Under additional assumptions, an instrumental variable approach identifies the causal effect of a given clincal decision.

To measure physician batch tendency, we use the physician's residualized leave-out average batch rate. This measure is derived from two steps following the approaches taken by @doyle2015measuring, @dobbie2018effects, and @eichmeyer2022pathways. First, we obtain residuals from a regression model, which includes all ED encounters in our sample period.

@@

Where $Y_i$ represents our main outcomes of interest: ED LOS, TIME TO RESULT, NUM IMAGES, 72HRA, and $X_i$ is the same as in the reduced-form approach. $Batched_i$ variable suffers from potential endogeneity concerns. For example, injury severity may be unobserved and correlated with need to run multiple tests, length of stay, and return with admission likelihood. Hence, we instrument $Batched_i$ with the assigned physician $j$'s underlying tendency to batch, $Tendency_{i,j}^{phys}$. We cluster robust standard errors at the physician level to account for the assignment process of patients to physicians.

### Identifying Assumptions

\
The reduced-form approach delivers an unbiased estimate of the causal effect of being treated by a higher tendency to batch physician, since assignment of patients to ED physicians is random, conditional on seasonality and shift (``conditional independence"). The residualization in equation (1) controls for more controls than required to achieve quasi-random assignment; they are included for statistical precision in measuring physician tendency to batch.

Our instrumental variable approach, which aims to recover the causal effect of having diagnostic tests batch ordered, relies on three additional assumptions: relevance, exclusion, and monotonicity. We reported a strong first stage (i.e., relevance) at the end of the previous Section. The exclusion restriction requires that the instrument must influence the outcome of interest only through its effect on test batching. This is perhaps our strongest assumption and is at its core, untestable. However, several features of the ED setting suggest that such violation may likely only have a small impact and may be less concerning than in other health care settings. First, unlike in primary care settings, where the patient and primary care provider have many repeat encounters, the scope of what the emergency physician can do to impact medium-term outcomes is limited and well-observed by the researcher. Second, any violation of the exclusion restriction needs to directly affect the specific outcome of interest. The channel by which ED physicians can influence length of stay relative outcomes is likely through testing and diagnosis. Nevertheless, we take this assumption seriously and perform a placebo check in Section [...] as well as various robustness checks in Section [...].

Finally, the monotonicity assumption is necessary for interpreting the coefficient estimates obtained from the IV approach as Local Average Treatment Effects (LATEs) if there are heterogeneous treatment effects. It requires that any patient who is (not) batched by a sequencer (batcher) would also (not) be batched by a batcher (sequencer) physician. The literature leveraging the judges design typically performs two informal tests for its implications. The first one provides that the first stage should be weakly positive for all subsamples (@dobbie2018effects). The second implication asserts that the instrument constructed by leaving out a particular subsample has predictive power over that same left-out subsample (@bhuller2020incarceration). Appendix Table  [...] presents both of these tests in the two columns for various subsamples of interest.

## Econometric Models

# Results and Discussion
## Determenants of Image Batching

@@

\small
\begin{equation}
Batched_{i,j} = \beta_0 + \beta_1 EXPERIENCE_{i,j} + \beta_2 SEXM_{i,j} + \beta_3 PATIENTS\_TBS + \beta_4 OCCUPANCY + X_{i,j} + \epsilon_i
\end{equation}

\normalsize
Where $Batched_{i,j}$ is a binary measure of whether physician $j$ batched tests for patient $i$. $EXPERIENCE_{i,j}$ is the number of years since the physician's residency graduation. $SEXM_{i,j}$ is a binary variable indicating the sex of the physician is male. $PATIENTS\_TBS$ is the number of remaining patients to be seen by the physician, excluding patient $i$, at the time of the patient's visit. $OCCUPANCY$ is the number of patients in the ED at the time of the arrival of the patient's visit. $X_{i,j}$ is the vectors of patient covariates described in the previous section. We cluster robust standard errors at the physician level.

@@

## Impact on Emergency Department Operations and Patient Outcomes

## Heterogeneity in Ordering Strategy

## Generealizability of Results

## Managerial Implications

## Robustness Checks

# Conclusion

## Contributions

## Conclusion

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Author template for INFORMS Journal on Data Science (ijds) [interim solution; new styles under construction]
%% Mirko Janc, Ph.D., INFORMS, mirko.janc@informs.org
%% ver. 0.91, March 2015 - updated November 2020 by Matthew Walls, matthew.walls@informs.org
%% Adapted for rticles by Rob J Hyndman Rob.Hyndman@monash.edu. Dec 2021
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[,mnsc,nonblindrev]{informs}

\OneAndAHalfSpacedXI
%%\OneAndAHalfSpacedXII % Current default line spacing
%%\DoubleSpacedXII
%%\DoubleSpacedXI

%% BEGIN MY ADDITIONS %%
\usepackage{hyperref}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{siunitx}
\usepackage{tablefootnote}
\usepackage{longtable}
\usepackage{threeparttable}
\usepackage{natbib}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{multirow}

%% END MY ADDITIONS %%

% Natbib setup for author-year style
\usepackage{natbib}
 \bibpunct[, ]{(}{)}{,}{a}{}{,}%
 \def\bibfont{\small}%
 \def\bibsep{\smallskipamount}%
 \def\bibhang{24pt}%
 \def\newblock{\ }%
 \def\BIBand{and}%

%% Setup of theorem styles. Outcomment only one.
%% Preferred default is the first option.
\TheoremsNumberedThrough     % Preferred (Theorem 1, Lemma 1, Theorem 2)
%\TheoremsNumberedByChapter  % (Theorem 1.1, Lema 1.1, Theorem 1.2)
\ECRepeatTheorems

%% Setup of the equation numbering system. Outcomment only one.
%% Preferred default is the first option.
\EquationsNumberedThrough    % Default: (1), (2), ...
%\EquationsNumberedBySection % (1.1), (1.2), ...

% For new submissions, leave this number blank.
% For revisions, input the manuscript number assigned by the on-line
% system along with a suffix ".Rx" where x is the revision number.
\MANUSCRIPTNO{}

%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%

% Outcomment only when entries are known. Otherwise leave as is and
%   default values will be used.
%\setcounter{page}{1}
%\VOLUME{00}%
%\NO{0}%
%\MONTH{Xxxxx}% (month or a similar seasonal id)
%\YEAR{0000}% e.g., 2005
%\FIRSTPAGE{000}%
%\LASTPAGE{000}%
%\SHORTYEAR{00}% shortened year (two-digit)
%\ISSUE{0000} %
%\LONGFIRSTPAGE{0001} %
%\DOI{10.1287/xxxx.0000.0000}%

% Author's names for the running heads
% Sample depending on the number of authors;
\RUNAUTHOR{%
Jameson, Soroush, Huckman
 and Hodgson
}
% \RUNAUTHOR{Jones and Wilson}
% \RUNAUTHOR{Jones, Miller, and Wilson}
% \RUNAUTHOR{Jones et al.} % for four or more authors
% Enter authors following the given pattern:
%\RUNAUTHOR{}

\RUNTITLE{Image Batching}

\TITLE{Image Batching: Batch Ordering Advanced Imaging Tests in the
Emergency Department}

\ARTICLEAUTHORS{%
\AUTHOR{Jacob Jameson}
\AFF{Harvard Kennedy School, Harvard
University, \EMAIL{\href{mailto:jacobjameson@g.harvard.edu}{\nolinkurl{jacobjameson@g.harvard.edu}}}}

\AUTHOR{Saghafian Soroush}
\AFF{Harvard Kennedy School, Harvard University, \EMAIL{}}

\AUTHOR{Robert Huckman}
\AFF{Harvard Business School, Harvard University, \EMAIL{}}

\AUTHOR{Nicole Hodgson}
\AFF{Mayo Clinic, Emergency Department, \EMAIL{}}

%
}

\ABSTRACT{Abstract goes here}

\KEYWORDS{3 to 6 keywords; that do not appear in the title}

\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Emergency departments (EDs) are high-pressure environments which
globally face escalating challenges such as overcrowding, resource
limitations, and increasing patient volumes
\citep[\citet{Sorup2013evaluation}]{mostafa2024strategies}. These
challenges can often be exacerbated by the need to provide timely and
accurate diagnoses to patients, which often requires the use of advanced
imaging tests such as computed tomography (CT) scans, ultrasound, X-ray,
and magnetic resonance imaging (MRI) \citep{waheed2022impact}. The
management of diagnostic testing is a critical component of operational
performance in EDs, where timely and accurate diagnoses can
significantly impact patient outcomes \citep{committee2015improving}. In
many EDs, however, physicians have considerable discretion over how they
order these tests, particularly with regard to advanced imaging
\citep{valtchinov2019use}. Yet little is known about the factors driving
physicians' decisions to exercise such discretion and how test ordering
should be managed when this discretion exists. In this paper, we explore
the causal effects of batch ordering advanced imaging tests, a practice
in which physicians order multiple tests simultaneously, on operational
performance and patient outcomes in the ED.

Worker discretion can improve system performance, but can sometimes
enable workers to ``choose the `wrong' task (operationally)''
\citep{vanDonselaar2010ordering}. We consider the physician's decision
to order imaging tests for their patient as an optimization problem,
where the physician must balance the trade-offs between the benefits of
ordering multiple tests simultaneously at the start of the patient
encounter, which may expedite the diagnostic process if the tests are
necessary for diagnosis and disposition, and the costs of ordering
multiple tests simultaneously, which may increase the total time spent
in the ED and the resource utilization of the department. Batching
stands in contrast to the more common practice of ordering tests
sequentially, where physicians order one test at a time, review the
results, and then decide whether to order additional tests. The decision
to batch order imaging tests is a form of discretion that physicians can
exercise in the ED.

@@

We investigate these questions using operational data from two large EDs
in the United States. We focus on imaging tests, which are commonly
ordered in the ED and are associated with significant resource
utilization. We use a unique dataset that contains detailed information
on the timing of test orders, the timing of test results, and the timing
of patient disposition. We exploit the random assignment of patients to
physicians to identify the causal effects of batch ordering imaging
tests on operational performance and patient outcomes. We find that
physicians deviate from the standard test ordering strategy 42\% of the
time. We show that physicians are more likely to batch order imaging
tests when they are more experienced, when there is an opportunity to
follow a superior test ordering strategy by deviating, and when there is
an opportunity to batch by deviating. When physicians deviate from the
standard test ordering strategy, their average time to disposition
increases by 13\%. Other performance dimensions, including resource
utilization and patient outcomes, are mostly unaffected. Our
calculations suggest that forgoing deviations would have led to faster
time to disposition, which could have saved 2,494 hours per year,
increasing annual profits by 3\%.

@@

\hypertarget{related-literature}{%
\section{Related Literature}\label{related-literature}}

\hypertarget{batch-ordering-advanced-imaging}{%
\section{Batch Ordering Advanced
Imaging}\label{batch-ordering-advanced-imaging}}

\hypertarget{drivers-of-image-batching}{%
\subsection{Drivers of Image Batching}\label{drivers-of-image-batching}}

\hypertarget{implications-for-emergency-department-operations-and-patient-outcomes}{%
\subsection{Implications for Emergency Department Operations and Patient
Outcomes}\label{implications-for-emergency-department-operations-and-patient-outcomes}}

\hypertarget{hetereogeneity-in-orderging-strategy}{%
\subsection{Hetereogeneity in Orderging
Strategy}\label{hetereogeneity-in-orderging-strategy}}

\hypertarget{setting-data-and-models}{%
\section{Setting, Data, and Models}\label{setting-data-and-models}}

\hypertarget{empirical-setting}{%
\subsection{Empirical Setting}\label{empirical-setting}}

@@

\hypertarget{data}{%
\subsection{Data}\label{data}}

Our primary dataset comes from the ED of the Mayo Clinic of Arizona, a
tertiary care hospital without obstetrical services, an inpatient
pediatrics unit, or a trauma designation. During the study period from
October 6, 2018, through December 31, 2019, the ED recorded 48,854
visits per year, managed across 26 treatment rooms and up to 9 hallway
spaces. The department is exclusively staffed by board-eligible or
board-certified emergency physicians (EPs), with rotating residents
overseeing about 10\% of patient volume. The data is summarized in Table
\ref{tab:summary_stats}.

@@

A key feature of the Mayo Clinic ED is its rotational patient assignment
system, in which patients arriving at the Mayo Clinic ED are randomly
assigned to physicians via a rotational patient assignment algorithm
\citet{traub2016emergency}, which removes potential selection bias
concerns for our analyses. In essence, barring arrival time and
shift-level variation, the physician-to-patient matching can be deemed
random. Table \ref{tab:wald_test} displays that patient encounters
(regarding chief complaints and emergency severity) are equitably
distributed across physicians within our study's cohort.

@@

Our research design focuses on adults who visit the Mayo Clinic of
Arizona ED. We observe \(48,854\) such visits during the study period.
To improve power, we drop encounters with rare ``reasons for visit''
(RVF) (\(<1000\) total encounters of this kind) and complaints where a
batch order occurs less than \(5\%\) of the time. Since batch orders are
rare for these cases, our physician batch tendency instrument could
suffer from a weak instrument problem if we included them. Examples of
complaints dropped include Skin Complaints and Urinary Complaints, as
well as other complaints where multiple modalities of imaging is
unlikely to occur. Excluding these conditions does not introduce
selection bias unless physician test batching tendency is orthogonal to
physician diagnosing behavior. While this assumption may be violated if
we were to use a very detailed level of chief complaint information upon
which to base our exclusion criterion, it is plausibly satisfied when
using broad complaint categories as we do. In order to estimate a
precise measure of physician-level batch tendency, we further restrict
our sample to the \(15,821\) encounters involving full-time physicians
who treat over 500 ED cases per year.

\hypertarget{instrumental-variable}{%
\subsubsection{Instrumental Variable}\label{instrumental-variable}}

\hfill\break
Our explanatory variable in the IV analysis, \(Batched_i\), is an
indicator for whether patient \(i\) has their tests batch ordered during
their ED encounter. We define ``batching'' in line with standard
emergency medicine practices. Batching occurs when a physician
simultaneously orders a comprehensive set of diagnostic tests, typically
covering a broad range of potential diagnoses. This contrasts with
sequential ordering, where tests are ordered sequentially based on the
information obtained from each test as needed.

@@

\hypertarget{dependent-variables}{%
\subsubsection{Dependent Variables}\label{dependent-variables}}

\hfill\break
\textit{Logarithm of Emergency Department Length of Stay (LOS)}: ED LOS
is a key performance metric for ED operations. It is defined as the time
from patient arrival to the ED to the time of discharge from the ED. We
use this metric to evaluate the impact of batch ordering on ED
efficiency. This variable is measured in minutes and in our sample is
right-skewed. We log-transform this variable to normalize its
distribution to meet the assumptions of linear regression.

@@

\textit{Total Imaging Tests Ordered}: Total imaging tests ordered
(NUM\_IMAGES) is a count of the number of distinct imaging tests ordered
for a patient during the ED encounter. We use this variable to evaluate
the impact of batch ordering on the number of imaging tests ordered.

@@

\hypertarget{identification-strategy}{%
\subsection{Identification Strategy}\label{identification-strategy}}

Our empirical strategy closely follows the literature that relies on
quasi-random assignment of agents to cases, often referred to as the
``judges design.'' Papers in this literature typically exploit variation
in the sentencing leniency of judges who work in the same court.
Similarly, we explore batching variation across physicians who work in
the same emergency department. In its reduced form, under the assumption
of quasi-random assignment, this approach allows researchers to identify
the causal effect of being assigned to different types of physicians
(i.e.~physiancs with a higher or lower tendency to batch order imaging).
Under additional assumptions, an instrumental variable approach
identifies the causal effect of a given clincal decision.

To measure physician batch tendency, we use the physician's residualized
leave-out average batch rate. This measure is derived from two steps
following the approaches taken by \citet{doyle2015measuring},
\citet{dobbie2018effects}, and \citet{eichmeyer2022pathways}. First, we
obtain residuals from a regression model, which includes all ED
encounters in our sample period.

@@

Where \(Y_i\) represents our main outcomes of interest: ED LOS, TIME TO
RESULT, NUM IMAGES, 72HRA, and \(X_i\) is the same as in the
reduced-form approach. \(Batched_i\) variable suffers from potential
endogeneity concerns. For example, injury severity may be unobserved and
correlated with need to run multiple tests, length of stay, and return
with admission likelihood. Hence, we instrument \(Batched_i\) with the
assigned physician \(j\)'s underlying tendency to batch,
\(Tendency_{i,j}^{phys}\). We cluster robust standard errors at the
physician level to account for the assignment process of patients to
physicians.

\hypertarget{identifying-assumptions}{%
\subsubsection{Identifying Assumptions}\label{identifying-assumptions}}

\hfill\break
The reduced-form approach delivers an unbiased estimate of the causal
effect of being treated by a higher tendency to batch physician, since
assignment of patients to ED physicians is random, conditional on
seasonality and shift (``conditional independence''). The
residualization in equation (1) controls for more controls than required
to achieve quasi-random assignment; they are included for statistical
precision in measuring physician tendency to batch.

Our instrumental variable approach, which aims to recover the causal
effect of having diagnostic tests batch ordered, relies on three
additional assumptions: relevance, exclusion, and monotonicity. We
reported a strong first stage (i.e., relevance) at the end of the
previous Section. The exclusion restriction requires that the instrument
must influence the outcome of interest only through its effect on test
batching. This is perhaps our strongest assumption and is at its core,
untestable. However, several features of the ED setting suggest that
such violation may likely only have a small impact and may be less
concerning than in other health care settings. First, unlike in primary
care settings, where the patient and primary care provider have many
repeat encounters, the scope of what the emergency physician can do to
impact medium-term outcomes is limited and well-observed by the
researcher. Second, any violation of the exclusion restriction needs to
directly affect the specific outcome of interest. The channel by which
ED physicians can influence length of stay relative outcomes is likely
through testing and diagnosis. Nevertheless, we take this assumption
seriously and perform a placebo check in Section {[}\ldots{]} as well as
various robustness checks in Section {[}\ldots{]}.

Finally, the monotonicity assumption is necessary for interpreting the
coefficient estimates obtained from the IV approach as Local Average
Treatment Effects (LATEs) if there are heterogeneous treatment effects.
It requires that any patient who is (not) batched by a sequencer
(batcher) would also (not) be batched by a batcher (sequencer)
physician. The literature leveraging the judges design typically
performs two informal tests for its implications. The first one provides
that the first stage should be weakly positive for all subsamples
(\citet{dobbie2018effects}). The second implication asserts that the
instrument constructed by leaving out a particular subsample has
predictive power over that same left-out subsample
(\citet{bhuller2020incarceration}). Appendix Table {[}\ldots{]} presents
both of these tests in the two columns for various subsamples of
interest.

\hypertarget{econometric-models}{%
\subsection{Econometric Models}\label{econometric-models}}

\hypertarget{results-and-discussion}{%
\section{Results and Discussion}\label{results-and-discussion}}

\hypertarget{determenants-of-image-batching}{%
\subsection{Determenants of Image
Batching}\label{determenants-of-image-batching}}

@@

\small

\begin{equation}
Batched_{i,j} = \beta_0 + \beta_1 EXPERIENCE_{i,j} + \beta_2 SEXM_{i,j} + \beta_3 PATIENTS\_TBS + \beta_4 OCCUPANCY + X_{i,j} + \epsilon_i
\end{equation}

\normalsize

Where \(Batched_{i,j}\) is a binary measure of whether physician \(j\)
batched tests for patient \(i\). \(EXPERIENCE_{i,j}\) is the number of
years since the physician's residency graduation. \(SEXM_{i,j}\) is a
binary variable indicating the sex of the physician is male.
\(PATIENTS\_TBS\) is the number of remaining patients to be seen by the
physician, excluding patient \(i\), at the time of the patient's visit.
\(OCCUPANCY\) is the number of patients in the ED at the time of the
arrival of the patient's visit. \(X_{i,j}\) is the vectors of patient
covariates described in the previous section. We cluster robust standard
errors at the physician level.

@@

\hypertarget{impact-on-emergency-department-operations-and-patient-outcomes}{%
\subsection{Impact on Emergency Department Operations and Patient
Outcomes}\label{impact-on-emergency-department-operations-and-patient-outcomes}}

\hypertarget{heterogeneity-in-ordering-strategy}{%
\subsection{Heterogeneity in Ordering
Strategy}\label{heterogeneity-in-ordering-strategy}}

\hypertarget{generealizability-of-results}{%
\subsection{Generealizability of
Results}\label{generealizability-of-results}}

\hypertarget{managerial-implications}{%
\subsection{Managerial Implications}\label{managerial-implications}}

\hypertarget{robustness-checks}{%
\subsection{Robustness Checks}\label{robustness-checks}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{contributions}{%
\subsection{Contributions}\label{contributions}}

\hypertarget{conclusion-1}{%
\subsection{Conclusion}\label{conclusion-1}}

@@

% Appendix here
% Options are (1) APPENDIX (with or without general title) or
%             (2) APPENDICES (if it has more than one unrelated sections)
% Outcomment the appropriate case if necessary
%
% \begin{APPENDIX}{<Title of the Appendix>}
% \end{APPENDIX}
%
%   or
%
% \begin{APPENDICES}
% \section{<Title of Section A>}
% \section{<Title of Section B>}
% etc
% \end{APPENDICES}

% Acknowledgments here
\ACKNOWLEDGMENT{The authors would like to thank \ldots{}}

\bibliographystyle{informs2014}
\bibliography{references.bib}

\end{document}
